---
title: "Simulating Rasters (Chapter 12)"
author: "Ed Rubin"
date: "2/15/2020"
output:
  html_document: default
  pdf_document: default
---

## Part 1: Simulate the data

In this simple tutorial, we simulate a raster data set containing $50 \times 50$ cells. 

```{r}
# Number of rows
  nx = 50
  ny = 50

# Set seed for reproducibility
  set.seed(123)
```

Next, we simulate random normal distributions and create three rasters.
```{r}
#Load
  pacman::p_load(raster, data.table, magrittr)

#Generate data
#Create combination of all x and y values
# Create a random normal distribution
  r1 = expand.grid(
    x = 1:nx,
    y = 1:ny
  ) %T>%
  setDT() %>%
  .[,val := rnorm(n = nx*ny, mean = x + y, sd = sqrt(ny))] %>%
  rasterFromXYZ()
  
  r2 = expand.grid(
    x = 1:nx,
    y = 1:ny
  ) %T>%
  setDT() %>%
  .[,val := rnorm(n = nx*ny, mean = (-0.5) * x + 2 * y, sd = sqrt(ny))] %>%
  rasterFromXYZ()
  
  r3 = expand.grid(
    x = 1:nx,
    y = 1:ny
  ) %T>%
  setDT() %>%
  .[,val := rnorm(n = nx*ny, mean = (1) * x - 1.25 * y, sd = sqrt(ny))] %>%
  rasterFromXYZ()
  
```

Also, construct a data frame with simulated population.

```{r}
  pop_r = expand.grid(
    x = 1:nx,
    y = 1:ny
  ) %T>%
  setDT() %>%
  .[,val := runif(n = nx*ny, min = 1, max = 100)] %>%
  rasterFromXYZ()
```

For each, we standardize the values.
```{r}
# Force to same range: 1 to 100
  r1 = r1 - cellStats(r1, min) + 1
  r1 = 100 * r1 / cellStats(r1, max)
  r2 = r2 - cellStats(r2, min) + 1
  r2 = 100 * r2 / cellStats(r2, max)
  r3 = r3 - cellStats(r3, min) + 1
  r3 = 100 * r3 / cellStats(r3, max)
```

Save out the data as a `Rda`.
```{r}
save(r1, r2, r3, pop_r, file = "raster_simulate.Rda")
```

##Part 2: Raster Math and Visuals

To check if the data is in good shape, we plot each raster.

```{r}
#Load graphing packages
  pacman::p_load(ggplot2, gridExtra, cowplot)

#Create the plots
  p1 = ggplot(data = as.data.frame(r1, xy = T)) +
    geom_raster(aes(x, y, fill = val)) +
    scale_fill_viridis_c("", option = "magma") +
    guides(fill = guide_colourbar(barwidth = 1, barheight = 15)) +
    coord_equal() +
    theme_void() +
    ggtitle("r1") + 
    theme(legend.position = "none", plot.title =  element_text(size=10, hjust = 0.5))
  
  p2 = ggplot(data = as.data.frame(r2, xy = T)) +
    geom_raster(aes(x, y, fill = val)) +
    scale_fill_viridis_c("", option = "magma") +
    guides(fill = guide_colourbar(barwidth = 1, barheight = 15)) +
    coord_equal() +
    theme_void() +
    ggtitle("r2") + 
  theme(legend.position = "none", plot.title =  element_text(size=10, hjust = 0.5))
  
  p3 = ggplot(data = as.data.frame(r3, xy = T)) +
    geom_raster(aes(x, y, fill = val)) +
    scale_fill_viridis_c("", option = "magma") +
    guides(fill = guide_colourbar(barwidth = 1, barheight = 15)) +
    coord_equal() +
    theme_void() + 
    ggtitle("r3") 
  plegend = p3 %>% get_legend()
  p3 = p3 + theme(legend.position = "none", plot.title =  element_text(size=10, hjust = 0.5))
  
  grid.arrange(p1, p2, p3, ncol = 3)
```