---
title: "Difference-in-Differences"
output: html_notebook
---

This notebook constructs the underlying simulated data for Chapter 8's discussion on Regression Discontinuity (RD). 

#Part 1: Create random norml distribution setting discontinuity at 0.

```{r}
#install.packages("pacman")
  pacman::p_load(rdrobust, ggplot2)

#Let's generate some data
  set.seed(3264)
  N <- 1e5  #Number of observations
  Z <- rnorm(N) #Running Variable
  Z0 <- 0            #Cut-off point
  Q <- Z > Z0

#Example data
  example_data <- data.frame(Z = Z,
                             Q = Q)

```

#Part 2: Simulate an outcome variable 

$$Y = \alpha + \zeta Z + \tau Q + \varepsilon$$

where $Y$ is the outcome, $Z$ is the running variable, $Q$ is the treatment indicator, and $\varepsilon$ is an identically and indepdently distributed noise parameter. The unknown object of interest is $\tau$ or the effect of changing one's treatment status on the outcome $Y$.

```{r}
#Set parameters
  varepsilon <- rnorm(N)
  tau <- 3
  alpha <- 0
  zeta <- 0.5
  
#Create outcome
  Y <-  alpha + zeta*Z + tau*Q + varepsilon
  
#Create data set
  simulated_rd_data <- data.frame(Y = Y,
                                  Z = Z,
                                  color = ifelse(Q == 1,'blue','red'))

```


